%% Alejandro Espinosa — DevOps & Platform Engineer — English CV v3 (Final)
%% Compile with: xelatex cv_english_final.tex
\documentclass[a4paper,9pt]{extarticle}

% ─── Packages ───
\usepackage[top=1.5cm, bottom=1.5cm, left=2.0cm, right=2.0cm]{geometry}
\usepackage{fontspec}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage[skip=0.2em]{parskip}
\usepackage{microtype}
\tolerance=1000
\hyphenpenalty=10000
\exhyphenpenalty=10000
\emergencystretch=1em
\usepackage{tikz}
\usepackage{fancyhdr}
\usepackage{lastpage}

% ─── Colors ───
\definecolor{headerband}{HTML}{1B2838}
\definecolor{accent}{HTML}{0D9488}
\definecolor{darktext}{HTML}{1A1A1A}
\definecolor{graytext}{HTML}{555555}
\definecolor{lightgray}{HTML}{D1D5DB}
\definecolor{linkcolor}{HTML}{0D7377}
\definecolor{headerwhite}{HTML}{F9FAFB}

% ─── Fonts ───{\small\color{graytext}\textit{Nov 2024 – Jun 2025: Relocation to Germany}}
\setmainfont{Poppins}[
  Path = ./fonts/Poppins/,
  UprightFont = Poppins-Regular,
  BoldFont = Poppins-Bold,
  ItalicFont = Poppins-Italic,
  BoldItalicFont = Poppins-BoldItalic,
]
\newfontfamily\lightfont{Poppins-Light}[
  Path = ./fonts/Poppins/,
  UprightFont = Poppins-Light,
]
\newfontfamily\mediumfont{Poppins-Medium}[
  Path = ./fonts/Poppins/,
  UprightFont = Poppins-Medium,
]

% ─── Hyperref ───
\hypersetup{
  colorlinks=true,
  urlcolor=linkcolor,
  linkcolor=linkcolor,
  pdfauthor={Alejandro Espinosa},
  pdftitle={Alejandro Espinosa — DevOps \& Platform Engineer},
}

% ─── Page style with footer on page 2+ ───
\fancypagestyle{cvfooter}{%
  \fancyhf{}%
  \renewcommand{\headrulewidth}{0pt}%
  \fancyfoot[C]{\small\color{graytext}Alejandro Espinosa\enspace\textcolor{lightgray}{|}\enspace Page \thepage}%
}
\fancypagestyle{firstpage}{%
  \fancyhf{}%
  \renewcommand{\headrulewidth}{0pt}%
}
\pagestyle{cvfooter}

% ─── Section formatting ───
\titleformat{\section}
  {\large\bfseries\color{accent}}
  {}
  {0em}
  {\MakeUppercase}
  [\vspace{-0.6em}{\color{accent}\rule{\textwidth}{1.2pt}}]

\titlespacing{\section}{0pt}{0.4em}{0.2em}

% ─── Custom commands ───
\newcommand{\jobentry}[4]{%
  {\mediumfont\color{darktext}#1}\hfill{\small\color{graytext}#4}\\
  {\color{accent}\textbf{#2}}\enspace{\small\color{graytext}| #3}\par\vspace{0.2em}
}

\newcommand{\jobentrywithdesc}[5]{%
  {\mediumfont\color{darktext}#1}\hfill{\small\color{graytext}#4}\\
  {\color{accent}\textbf{#2}}\enspace{\small\color{graytext}| #3}\\[0.15em]
  {\small\color{graytext}\textit{#5}}\par\vspace{0.2em}
}

\newcommand{\earlierrole}[4]{%
  {\small\mediumfont\color{darktext}#1} {\small\color{graytext}| #2 | #3}\\
  {\small\color{graytext}#4}\par\vspace{0.3em}
}

\newcommand{\skillcat}[2]{%
  {\mediumfont\small\color{accent}#1:}\enspace{\small #2}\par\vspace{0.15em}
}

% ─── List formatting ───
\setlist[itemize]{
  leftmargin=3em,
  itemsep=0.1em,
  parsep=0pt,
  topsep=0.15em,
  label={\textcolor{accent}{\textbullet}},
}

% ─── Document ───
\begin{document}
\color{darktext}
\thispagestyle{firstpage}

% ══════════════════════════════════════════════════════════════
%   HEADER BAND
% ══════════════════════════════════════════════════════════════

\noindent
\begin{tikzpicture}[remember picture, overlay]
  \fill[headerband]
    ([yshift=0.5cm]current page.north west) rectangle
    ([yshift=-3.6cm]current page.north east);
\end{tikzpicture}%
%
\vspace*{-1.0cm}
\begin{center}
  {\LARGE\bfseries\color{headerwhite}ALEJANDRO ESPINOSA}\\[0.3em]
  {\large\lightfont\color{accent}DevOps \& Platform Engineer}\\[0.55em]
  {\small\color{headerwhite}%
    Bochum, Germany\enspace{\color{accent}|}\enspace
    \href{mailto:workwithme@alejoespinosa.com}{\color{headerwhite}workwithme@alejoespinosa.com}\enspace{\color{accent}|}\enspace
    +49\,162\,1050507\\[0.2em]
    \href{https://www.linkedin.com/in/alejoespinosa}{\color{headerwhite}linkedin.com/in/alejoespinosa}\enspace{\color{accent}|}\enspace
    \href{https://github.com/halbardhobby}{\color{headerwhite}github.com/halbardhobby}
  }\\[0.45em]
  {\small\mediumfont\color{accent}EU Blue Card\enspace{\color{headerwhite}|}\enspace
  \color{accent}Immediately Available\enspace{\color{headerwhite}|}\enspace
  \color{accent}No Sponsorship Required}
\end{center}

\vspace{0.3cm}

% ══════════════════════════════════════════════════════════════
%   SUMMARY
% ══════════════════════════════════════════════════════════════

\section{Summary}

DevOps and Platform Engineer with 10 years of experience building self-service platform engineering foundations and automating cloud infrastructure across AWS, GCP, and Azure. Delivered enterprise-scale infrastructure for HBO Max, Disney Parks, and clients in healthcare and fintech. Built Kubernetes platforms from scratch in both startup and enterprise environments. Led cloud migration for a 1,000-employee firm with 40\%+ cost reduction. Former university lecturer with a focus on developer experience, self-service tooling, and comprehensive documentation.

% ══════════════════════════════════════════════════════════════
%   SKILLS (immediately after summary — tech hiring scans here)
% ══════════════════════════════════════════════════════════════
\section{Skills}

\skillcat{Platform Engineering \& IaC}{Terraform (certified), Terragrunt, Ansible}
\skillcat{Containers \& Orchestration}{Kubernetes, Docker, Helm}
\skillcat{CI/CD \& GitOps}{GitHub Actions, ArgoCD, Jenkins, AWS CodePipeline, GCP Cloud Build}
\skillcat{Cloud Platforms}{AWS, GCP, Azure}
\skillcat{Observability \& Reliability}{Prometheus, Grafana, CloudWatch, Datadog (exposure), SLO/SLI monitoring, Alerting}
\skillcat{Security \& Compliance}{Container security, RBAC, Secrets management, GDPR-aware infrastructure}
\skillcat{Programming}{Python, Go, Bash, JavaScript, SQL}
\skillcat{Systems}{Linux (primary OS), Systemd, DNS, Load balancing}
\skillcat{AI Tooling}{LLM-assisted workflows (Claude, Gemini), custom AI agent development}

% ══════════════════════════════════════════════════════════════
%   EXPERIENCE
% ══════════════════════════════════════════════════════════════

\section{Work Experience}

\jobentrywithdesc{DevOps Engineer}{sovity GmbH}{Dortmund, Germany}{Jul 2025 -- Dec 2025}{Built CI/CD and platform engineering foundations for a data sovereignty startup on Azure and Kubernetes.}
\begin{itemize}
  \item Deployed core product hub into Kubernetes across 2 environments (dev + prod) using GitHub Actions, Terraform, and Terragrunt on Azure, with RBAC and secrets management configured.
  \item Built self-service deployment pipelines enabling developers to deploy with all infrastructure provisioned automatically --- reducing deployment lead time from hours to minutes.
  \item Reduced infrastructure tech debt by auditing and eliminating redundant dependencies and unused IaC modules across the Terragrunt codebase.
\end{itemize}
\vspace{1em}

\jobentry{Senior DevOps Engineer}{Globant}{Bogot\'{a}, Colombia}{May 2021 -- Nov 2024}
{\small\color{graytext}\textit{Deployed as senior consultant to multiple enterprise clients, each with independent stacks, teams, and SLA requirements.}}

\vspace{0.2em}
{\small\mediumfont\color{accent}HBO Max}
\begin{itemize}
  \item Automated service recovery and incident response workflows on AWS for a streaming platform serving 70M+ subscribers, reducing manual intervention and improving mean time to recovery across production services.
  \item Participated in on-call rotations and authored operational runbooks for a 10-person DevOps team.
\end{itemize}

\vspace{0.15em}
{\small\mediumfont\color{accent}Disney Parks}
\begin{itemize}
  \item Managed daily operational data synchronization pipelines across Disney's US parks (incl.\ Galaxy's Edge), feeding guest-facing apps and ride operations systems with zero-downtime requirements.
  \item Performed data validation and quality checks across sync outputs, including on-demand production data adjustments to maintain operational accuracy for downstream park systems.
\end{itemize}

\vspace{0.15em}
{\small\mediumfont\color{accent}Major Fintech Platform (NDA)}
\begin{itemize}
  \item Implemented GitOps deployment workflows using ArgoCD on Kubernetes, enabling declarative infrastructure management and automated rollbacks in production.
  \item Reviewed infrastructure-as-code PRs and implemented automated testing for IaC changes to reduce deployment risk.
\end{itemize}

\vspace{0.15em}
{\small\mediumfont\color{accent}Healthcare / MedTech Platform}
\begin{itemize}
  \item Implemented preemptive auto-scaling via queue depth monitoring for a medical imaging pipeline processing 3D models (\raise.17ex\hbox{$\scriptstyle\sim$}12\,min/request), ensuring SLA compliance and uninterrupted patient treatment workflows.
  \item Managed Terraform state and conducted cost optimization across AWS resources in a 5-person DevOps team.
\end{itemize}

\vspace{1em}
\newpage

\jobentry{University Lecturer}{Universidad del Rosario}{Bogot\'{a}, Colombia}{Sep 2019 -- Jan 2021}
\begin{itemize}
  \item Created ``Introduction to Industry 4.0'' for law students covering digital evidence, legal edge cases, and early GPT applications --- bridging technical concepts for non-technical audiences.
  \item Delivered courses for Colombian government institutions: Python for non-technical staff, DevOps for developers, and microservices architecture for middle management. Over 150 students total.
\end{itemize}

\vspace{1em}

\jobentrywithdesc{Backend Developer \& DevOps Engineer}{Efecto Studios}{Bogot\'{a}, Colombia}{Jun 2018 -- Jun 2019}{Introduced CI/CD to a game studio with no prior DevOps culture --- a first in the Colombian game industry.}
\begin{itemize}
  \item Built CI pipeline using Jenkins, Unreal Engine 4, and AWS that enabled daily feature delivery, overhauling the studio's entire release workflow.
  \item Integrated telemetry for Decoherence (released on Steam), enabling data-driven balance patches post-launch.
\end{itemize}

\vspace{1em}

\jobentry{Research Assistant — Interactive Media Lab (COLIVRI)}{Universidad de los Andes}{Bogot\'{a}, Colombia}{Sep 2017 -- Feb 2018}
\begin{itemize}
  \item Developed experimental VR applications using Unity and C\# to test hypotheses around visual perception in immersive environments, supporting the research team's experimental framework.
\end{itemize}

\vspace{1em}

\jobentrywithdesc{Innovation Engineer — Cloud Migration \& Monitoring}{INGETEC S.A.}{Bogot\'{a}, Colombia (during M.S.\ studies)}{Mar 2017 -- Aug 2017}{Full-time cloud transformation role at one of Colombia's largest civil engineering consultancies (\raise.17ex\hbox{$\scriptstyle\sim$}1,000 employees). Left due to scheduling conflicts with Master's program.}
\begin{itemize}
  \item Led a 3-person team to migrate on-premises infrastructure to Google Cloud Platform, moving engineering workloads and historical project data to the cloud.
  \item Achieved 40\%+ reduction in technology and maintenance costs through the migration.
  \item Drove cloud adoption strategy within a lean IT team of 6, managing stakeholder communication across technical and non-technical departments.
\end{itemize}

\vspace{1em}

% ── Earlier Roles ──
{\mediumfont\color{accent}\small EARLIER ROLES}

\earlierrole{Analyst Software Engineer}{We\{co\}}{Aug 2016 -- Jan 2017}{Built serverless backend on Google Cloud Functions and Firebase. Full development lifecycle on GCP.}
\earlierrole{Research Assistant — Industrial Engineering}{Universidad de los Andes}{Jan 2016 -- Nov 2016}{Simulation models for knowledge distribution and inter-team coordination strategies.}
\earlierrole{Backend Developer}{On3D Studios}{Jul 2015 -- Oct 2015}{Early career game/web dev. Contributed to Quantum Replica (later released on Steam).}
{\small\mediumfont\color{darktext}Relocation to Germany} {\small\color{graytext}| Nov 2024 -- Jun 2025}\\

% ══════════════════════════════════════════════════════════════
%   CERTIFICATIONS
% ══════════════════════════════════════════════════════════════

\section{Certifications}

{\mediumfont HashiCorp Terraform Associate}\hfill{\small\color{graytext}Dec 2021}\\
{\mediumfont IELTS Academic 7.5 (CEFR C1)}\hfill{\small\color{graytext}Sep 2023}

% ══════════════════════════════════════════════════════════════
%   EDUCATION
% ══════════════════════════════════════════════════════════════

\section{Education}

{\mediumfont M.S.\ Computer Science}\enspace{\small\color{graytext}| Universidad de los Andes, Colombia | 2016--2019}\\
{\mediumfont B.S.\ Computer Science}\enspace{\small\color{graytext}| Universidad de los Andes, Colombia | 2010--2015 (incl.\ Japanese minor)}

% ══════════════════════════════════════════════════════════════
%   LANGUAGES
% ══════════════════════════════════════════════════════════════

\section{Languages}

{\small
\begin{tabularx}{\textwidth}{@{}l X l X@{}}
  \mediumfont Spanish & Native & \mediumfont German & B1 \\
  \mediumfont English & C1 (IELTS Academic 7.5) & \mediumfont Japanese & JLPT N3 
\end{tabularx}
}

% ══════════════════════════════════════════════════════════════
%   PROJECTS
% ══════════════════════════════════════════════════════════════

\section{Projects}

{\mediumfont Home Lab} --- Bare-metal Ubuntu server running containerized services (Prometheus, Grafana, Pi-hole DNS, Gitea, media stack, reverse proxy) deployed via GitOps workflows. Infrastructure version-controlled in GitHub.

\end{document}
